---
title: "Simulation Within-Subject Mediation"
author: "Yenny Webb-Vargas"
date: "Wednesday, March 18, 2015"
output: html_document
---

- I simulate scalar mediation with confounding across people
- I try two bootstraps:
* perform analysis within person, then boostrap causal effect
* do bootstrap for each person, then sample from the bootstrap distribution

Sampling from bootstrap distribution
```{r}
N = 20 
J = 48

# Parameters
beta = c(2,3,1.5)
sd_x = 1

# Parameters within a person
sd_x_wd = 0.5

set.seed(493240)
beta_ws = t(sapply(1:N, function(i) rnorm(n=3, sd=sd_x_wd)))
sd_y_wd = 3


# Simulating data
x = matrix(rnorm(N*J, sd=sd_x), ncol=N)
z = sapply(1:N, function(n) sample(c(rep(1,J/2), rep(0,J/2)), replace=FALSE))
y = sapply(1:N, function(n) cbind(1, x[,n],z[,n])%*%(beta_ws[n,]+beta) + rnorm(J, sd=sd_y_wd))
plot(x,y)
```

```{r}
hat_beta = sapply(1:N, function(n) lm(y[,n]~x[,n]+z[,n])$coeff)
ATE = mean(hat_beta[grep("z",rownames(hat_beta)),])
hist(hat_beta[3,])
```

## Bootstrap people
```{r}
B=1000
myboot_sample = sapply(1:B, function(b) sample(1:N, replace=TRUE))
boot_ATE = sapply(1:B, function(b) c(mean(hat_beta[3,myboot_sample[,b]]),sd(hat_beta[3,myboot_sample[,b]])^2))
quantile(boot_ATE[1,], c(0.025, 0.975))
c(mean(boot_ATE[1,]),sd(boot_ATE[1,]))
hist(boot_ATE[1,])
```

## Combine using multiple imputation methods
```{r}
library(mitools)

list_beta = list()
for(n in 1:N){
  list_beta[[n]] = lm(y[,n]~x[,n]+z[,n])
}
(MI_beta = MIcombine(list_beta))
```

# Simulation
```{r}
S = 1500
boot_results = matrix(NA,ncol=4,nrow=S)
```


```{r, cache=TRUE}
set.seed(810014)
for(s in 1:S){
  N = 20 
  J = 48
  
  # Parameters
  beta = c(2,3,1.5)
  sd_x = 1
  
  # Parameters within a person
  sd_x_wd = 0.5
  
  beta_ws = t(sapply(1:N, function(i) rnorm(n=3, sd=sd_x_wd)))
  sd_y_wd = 3
  
  
  # Simulating data
  x = matrix(rnorm(N*J, sd=sd_x), ncol=N)
  z = sapply(1:N, function(n) sample(c(rep(1,J/2), rep(0,J/2)), replace=FALSE))
  y = sapply(1:N, function(n) cbind(1, x[,n],z[,n])%*%(beta_ws[n,]+beta) + rnorm(J, sd=sd_y_wd))
  
  # within-subject analysis
  hat_beta = sapply(1:N, function(n) lm(y[,n]~x[,n]+z[,n])$coeff)
  ATE = mean(hat_beta[grep("z",rownames(hat_beta)),])
  
  # Bootstrap Inference
  B=1000
  boot_sample = sapply(1:B, function(b) sample(1:N, replace=TRUE))
  boot_ATE = sapply(1:B, function(b) mean(hat_beta[3,boot_sample[,b]]))
  boot_results[s,] = c(quantile(boot_ATE, c(0.025, 0.975)), mean(boot_ATE),sd(boot_ATE))
}
```

```{r}
apply(boot_results[,3:4], 2,mean)
mean(boot_results[,1] < beta[3] & boot_results[,2] > beta[3]) 
boot_CI = boot_results[,3] + boot_results[,4]%*%t(qnorm(c(0.025, 0.975)))
mean(boot_CI[,1] < beta[3] & boot_CI[,2] > beta[3]) 
```

# Simulation with samples from bootstrap samples

```{r}
set.seed(493240)

N = 20 
J = 48

# Parameters
beta = c(2,3,1.5)
sd_x = 1

# Parameters within a person
sd_x_wd = 0.5

beta_ws = t(sapply(1:N, function(i) rnorm(n=3, sd=sd_x_wd)))
sd_y_wd = 3


# Simulating data
x = matrix(rnorm(N*J, sd=sd_x), ncol=N)
z = sapply(1:N, function(n) sample(c(rep(1,J/2), rep(0,J/2)), replace=FALSE))
y = sapply(1:N, function(n) cbind(1, x[,n],z[,n])%*%(beta_ws[n,]+beta) + rnorm(J, sd=sd_y_wd))
```

## boostrap within people

```{r}
library(boot)
set.seed(4174)

model = function(dta,index){
  dta_b = dta[index,]
  result = lm(y~x+z, data=dta_b)
  return(result$coef)
}
```

```{r}
ws_boot = list()

for(n in 1:N){
  dta = data.frame(x = x[,n], y = y[,n], z = z[,n])
  ws_boot[[n]] = boot(dta, statistic=model,R=B)
} 

B_n = 1500
boot_sample = sapply(1:N,FUN = function(n) sample(x =ws_boot[[n]]$t[,3], size = B_n, replace=TRUE))
boot_mean = apply(boot_sample, 1, mean)
mean(boot_mean)
quantile(boot_mean, c(0.025, 0.975))
```

## Simulation

```{r}
S = 1000
boot_results_ws = matrix(NA,ncol=4,nrow=S)
```


```{r, cache=FALSE}
set.seed(2935959)

for(s in 1:S){
  N = 20 
  J = 48
  
  # Parameters
  beta = c(2,3,1.5)
  sd_x = 1
  
  # Parameters within a person
  sd_x_wd = 0.5
  
  beta_ws = t(sapply(1:N, function(i) rnorm(n=3, sd=sd_x_wd)))
  sd_y_wd = 3
  
  
  # Simulating data
  x = matrix(rnorm(N*J, sd=sd_x), ncol=N)
  z = sapply(1:N, function(n) sample(c(rep(1,J/2), rep(0,J/2)), replace=FALSE))
  y = sapply(1:N, function(n) cbind(1, x[,n],z[,n])%*%(beta_ws[n,]+beta) + rnorm(J, sd =sd_y_wd))
  
  # Analysis
  ws_boot = list()
  
  for(n in 1:N){
    dta = data.frame(x = x[,n], y = y[,n], z = z[,n])
    ws_boot[[n]] = boot(dta, statistic=model,R=B)
  } 
  
  B_n = 1500
  boot_sample = sapply(1:N,FUN = function(n) sample(x =ws_boot[[n]]$t[,3], size = B_n, replace=TRUE))
  boot_mean = apply(boot_sample, 1, mean)
  boot_results_ws[s,] = c(quantile(boot_mean, c(0.025, 0.975)),mean(boot_mean), sd(boot_mean))
}
save(boot_results_ws, file='scalar-lm-simulation-boot_results.Rdata')
```

```{r}
load('scalar-lm-simulation-boot_results.Rdata')
apply(boot_results_ws[,3:4], 2,mean)
mean(boot_results_ws[,1] < beta[3] & boot_results_ws[,2] > beta[3]) 
boot_CI_ws = boot_results_ws[,3] + boot_results_ws[,4]%*%t(qnorm(c(0.025, 0.975)))
mean(boot_CI_ws[,1] < beta[3] & boot_CI_ws[,2] > beta[3]) 
```
